Effects of bariatric surgery and weight loss on resting state functional connectivity

### 2020-12-18 manuscript/draft_rsfmri.Rmd
FC result tables; two design options

# [ ] Evaluate SPM model comparison (more activation in SPM..)
# [ ] Discussion
# [ ] hervorheben, was von Präregistrierung abweicht
# [ ] brain images mit nilearn plotting tools einbinden
# [ ] Upload auf Neurovolt, cf. cobidas reporting guidelines
# [ ] re-do analysis with new sample
# [ ] repeat analysis with exclusion of 10% of worst data
# [ ] Network tests

@@

```{r tableSample, echo=FALSE}
tab_sample <-
  knitr::kable(
    tab_sample,
    row.names = TRUE,
    format = "latex",
    booktabs = T,
    caption = "Distribution of data points at months after intervention")
tab_sample
```

@@

tab_descr <-
  knitr::kable(
    descr,
    row.names = FALSE,
    format = "latex",
    booktabs = T,
    caption = "BMI and mean FD of participants acquired in each condition and time point")

@@

## Quality control
We performed different quality controls for the rsfMRI data.
As described in @Beyer_2020, there was a significant interaction of group and time on mFD. In particular, the intervention group reduced in head motion in the follow-up assessments compared to the control group.
We also inspected DVARS, a measurement of whole-brain variation in signal which is often reflective of noise in the data. We saw that DVARS did not qualitatively differ between groups and timepoints, and that the correlation of DVARS and mFD remained similar over time points.
This indicates that even though there was a decrease in head motion in the intervention versus control group, the impact of head motion on imaging data remained similar over the study period.
```{r tableQA, fig.cap="Summary table of QA measures"}
for (i in 1:nrow(final)) {
  tmp = strsplit(toString(final[i, "scan_dir"]), '/')[[1]][8]
  conf = read.csv(
    paste0("/data/pt_02161/preprocessed/resting/detailedQA/metrics/",tmp,"/confounds.csv")
  )
  tmp = cor.test(conf$DVARS, conf$FD)
  final[i, "corr_DVARS_FD"] = tmp$estimate
  final[i, "mean_DVARS"] = mean(conf$DVARS, na.rm = TRUE)
  final[i, "max_DVARS"] = max(conf$DVARS, na.rm = TRUE)
  final[i, "mean_outliers"] = mean(conf$outliers, na.rm = TRUE)
}

@@

tab_qc_fc <-
  knitr::kable(
    round(res,3),
    col.names = colnames(res),
    row.names = TRUE,
    format = "latex",
    booktabs = T,
    caption = "Summary of quality metrics for different denoising pipelines across conditions and time points")

@@

#### with adjustment for age and sex
Similarly, there were only non-findings when reducing the number of nuisance covariates by leaving out mean FD, no connectivity was found between the PCC with one other voxel.
Results did not differ between analyses of connetivity maps from different preprocessing pipelines.
<!-- There was no other significant co-activation for any other region of interest for neither of the preprocessing variants. -->

@@

## Effect of mean and change in BMI (exploratory analysis)

>>>>

# [ ] Evaluate SPM model comparison (more activation in SPM..)
# [ ] Discussion
# [ ] brain images mit nilearn plotting tools einbinden
# [ ] Upload auf Neurovolt, cf. cobidas reporting guidelines
# [ ] re-do analysis with new sample
# [ ] repeat analysis with exclusion of 10% of worst data
# [ ] Network tests

@@

```{r tableSample, echo=FALSE}
tab_sample <-
  knitr::kable(
    tab_sample,
    row.names = TRUE,
    format = "latex",
    booktabs = T,
    linesep = "",
    caption = "Distribution of data points at months after intervention")
tab_sample
```

@@

tab_descr <-
  knitr::kable(
    descr,
    row.names = FALSE,
    format = "latex",
    booktabs = T,
    linesep = "",
    caption = "BMI and mean FD of participants acquired in each condition and time point")

@@

## Quality control
We performed different quality controls for the rsfMRI data.
As described in @Beyer_2020, there was a significant interaction of group and time on mFD. In particular, the intervention group reduced in head motion in the follow-up assessments compared to the control group.
We also inspected DVARS, a measurement of whole-brain variation in signal which is often reflective of noise in the data. We saw that DVARS did not qualitatively differ between groups and time points, and that the correlation of DVARS and mFD remained similar over time points.
This indicates that even though there was a decrease in head motion in the intervention versus control group, the impact of head motion on imaging data remained similar over the study period.
```{r tableQA, fig.cap="Summary table of QA measures"}

for (i in 1:nrow(final)) {
  tmp = strsplit(toString(final[i, "scan_dir"]), '/')[[1]][8]
  conf = read.csv(
    paste0("/data/pt_02161/preprocessed/resting/detailedQA/metrics/",tmp,"/confounds.csv")
  )
  tmp = cor.test(conf$DVARS, conf$FD)
  final[i, "corr_DVARS_FD"] = tmp$estimate
  final[i, "mean_DVARS"] = mean(conf$DVARS, na.rm = TRUE)
  final[i, "max_DVARS"] = max(conf$DVARS, na.rm = TRUE)
  final[i, "mean_outliers"] = mean(conf$outliers, na.rm = TRUE)
}
```

@@

tab_qc_fc <-
  knitr::kable(
    round(res,3),
    col.names = colnames(res),
    row.names = TRUE,
    format = "latex",
    booktabs = T,
    linesep = "",
    caption = "Summary of quality metrics for different denoising pipelines across conditions and time points")

@@

#### with adjustment for age and sex
Similarly, there were only non-findings when reducing the number of nuisance covariates by leaving out mean FD. Results did not differ between analyses of connetivity maps from different preprocessing pipelines.
<!-- There was no other significant co-activation for any other region of interest for neither of the preprocessing variants. -->

@@

## Effect of mean and change in BMI (exploratory analysis)
```{r tableResBMI, include=FALSE}
result_dir <- "/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis"
BMImodel_dirs <- 
  c("noExclFD/BMI_total/brain/PCC_cc_z/bmi-age-sex-meanFD_WB-c01cl",# deact
    "noExclFD/BMI_total/brain/PCC_cc_z/bmi-age-sex_WB-c01cl")# deact
FDBMImodel_dirs <- 
  c("noExclFD/FD_total/brain/Nacc_cc_z/bmi-fd-age-sex_WB-c02cl", # deact
    "noExclFD/FD_total/brain/Nacc_cc_z/bmi-fd-age-sex_WB-c03cl", # act
    "noExclFD/FD_total/brain/Nacc_gsr_z/bmi-fd-age-sex_WB-c03cl", # act
    "noExclFD/FD_total/brain/PCC_cc_z/bmi-fd-age-sex_WB-c01cl") # deact
FDmodel_dirs <- 
  c("noExclFD/FD_total/brain/Nacc_cc_z/fd-age-sex_WB-c01cl", # act
    "noExclFD/FD_total/brain/Nacc_gsr_z/fd-age-sex_WB-c01cl") # act
deact_file <- "results_deact.csv"
act_file <- "results_act.csv"
```

```{r}
# noExclFD, total sample, all tp
### first option
# BMIagesexfd_PCCcc_avgBMI_deact
df1 <- read.csv(file.path(result_dir,BMImodel_dirs[1],deact_file))
colnr <- ncol(df1)
df1$seed <- NULL
df1 <- rbind(rep(NA, colnr),df1)
df1[1:2,"seed"] <- c("\\textit{average BMI (decrease)}","PCC (cc)")
df1$covariates <- NA
df1$covariates[2] <- "age, sex, log mFD"
# BMIagesex_PCCcc_avgBMI_deact 
df2 <- read.csv(file.path(result_dir,BMImodel_dirs[2],deact_file))
df2$seed <- NULL
df2 <- rbind(rep(NA, colnr),df2)
df2[1:2,"seed"] <- c("\\textit{average BMI (decrease)}","PCC (cc)")
df2$covariates <- NA
df2$covariates[2] <- "age, sex"

### alternative (with group_row())
df1 <- read.csv(file.path(result_dir,BMImodel_dirs[1],deact_file))
colnr <- ncol(df1)
df1$seed <- NULL
df1[1,"seed"] <- c("PCC (cc)")
df1$covariates <- NA
df1$covariates[1] <- "age, sex, log mFD"
# BMIagesex_PCCcc_avgBMI_deact 
df2 <- read.csv(file.path(result_dir,BMImodel_dirs[2],deact_file))
df2$seed <- NULL
df2[1,"seed"] <- c("PCC (cc)")
df2$covariates <- NA
df2$covariates[1] <- "age, sex"

# BMIFDagesex_NACCcc_cggBMI_deact 
df3 <- read.csv(file.path(result_dir,FDBMImodel_dirs[1],deact_file))
df3$seed <- NA
 df3 <- rbind(rep(NA, colnr),df3)
df3[1:2,"seed"] <- c("\\textit{change in BMI (decrease)}","NAcc (cc)")
df3$covariates <- NA
df3$covariates[2] <- "age, sex"
# BMIFDagesex_NACCcc_avgFD_act
 df4 <- read.csv(file.path(result_dir,FDBMImodel_dirs[2],act_file))
df4$seed <- NA
df4 <- rbind(rep(NA, colnr),df4)
df4[1:2,"seed"] <- c("\\textit{average log mean FD (increase)}","NAcc (cc)")
df4$covariates <- "age, sex"
# BMIFDagesex_NACCgsr_avgFD_act 
df5 <- read.csv(file.path(result_dir,FDBMImodel_dirs[3],act_file))
df5$seed <- NA
df5 <- rbind(rep(NA, colnr),df5)
df5[1:2,"seed"] <- c("\\textit{average log mean FD (increase)}","NAcc (gsr)")
df5$covariates <- NA
df5$covariates[2] <- "age, sex"
# BMIFDagesex_PCCcc_avgBMI-deact 
df6 <- read.csv(file.path(result_dir,FDBMImodel_dirs[4],deact_file))
df6$seed <- NA
df6 <- rbind(rep(NA, colnr),df6)
df6[1:2,"seed"] <- c("\\textit{average BMI (decrease)}","PCC (cc)")
df6$covariates <- NA
df6$covariates[2] <- "age, sex"

# BMIFDagesex_NACCcc_avgFD_act 
df7 <- read.csv(file.path(result_dir,FDmodel_dirs[1],act_file))
df7$seed <- NA
df7 <- rbind(rep(NA, colnr),df7)
df7[1:2,"seed"] <- c("\\textit{average log mean FD (increase)}","NAcc (cc)")
df7$covariates <- NA
df7$covariates[2] <- "age,sex"
# BMIFDagesex_NACCgsr_avgFD_act <
df8 <- read.csv(file.path(result_dir,FDmodel_dirs[2],act_file))
df8$seed <- NA
df8 <- rbind(rep(NA, colnr),df8)
df8[1:2,"seed"] <- c("\\textit{average log mean FD (increase)}", "NAcc (gsr)")
df8$covariates <- NA
df8$covariates[2] <- "age, sex"

BMImodel <- rbind(df1,df2)
FDmodel <- rbind(df7,df8)
BMIFDmodel <- rbind(df3,df4,df5,df6)
```

```{r include = FALSE}
model_list <- list(BMImodel,FDmodel,BMIFDmodel)

for (i in 1:length(model_list)){
  model_list[[i]]$anat <- NA
  model_list[[i]]$hemisphere <- NA
  # select columns
  model_list[[i]] <- model_list[[i]][,c("seed","covariates","anat","hemisphere","cluster_p.FWE.corr.","cluster_equivk","cluster_equivkZ","peak_Z","X_x","y","z..mm.")]
  # name columns
  colnames(model_list[[i]]) <- c("Seed","Covariates","Anatomical region","L/R","clusterwise FWE-corr. P","cluster size","Z score of cluster size","Z score in local maximum","X","Y","Z") 
}

for (i in 1:length(model_list)){
  # only keep cluster statistics and peak local maximum
  model_list[[i]] <- model_list[[i]][!is.na(model_list[[i]]$Seed) | !is.na(model_list[[i]]$`clusterwise FWE-corr. P`),]
}

BMImodel <- model_list[[1]]
FDmodel <- model_list[[2]]
BMIFDmodel <- model_list[[3]]
#rm(df1,df2,df3,df4,df5,df6,df7,df8)

```{r tabModel}
# hide NA in tables
fn1 <- "R, right hemisphere;L, left hemisphere; MNI (Montreal Neurological Institute) coordinates of primary peak location; x = sagittal, y = coronal,"
fn2 <- "z = axial. To identify significant clusters, we applied a cluster size threshold with p < 0.001 determined by Wild Bootstrap of 1000 samples."
title_vec <- c("BMI", "FD", "BMI-FD")

table_list <- list()
for (i in 1:length(model_list)){
  table_list[[i]] <-
    model_list[[i]] %>%
    knitr::kable(
      escape = FALSE,     # use font spec of latex with kableExtra
      col.names = colnames(model_list[[i]]),
      row.names = FALSE,
      format = "latex",
      booktabs = T,
      linesep = "",      # disable "\\addlinespace" at every 5th line
      caption = sprintf("Changes in functional connectivity in whole brain analysis for %s model",title_vec[i])
    ) %>%
    kable_styling(latex_options = c('hold_position')) %>% # "scale_down"
    column_spec(c(3, 5, 7, 8), width = "2.4cm") %>%
    add_footnote(c(fn1, fn2), notation = "none") %>%
    kableExtra::landscape()
}
tab_BMImodel <- table_list[[1]]
tab_FDmodel <- table_list[[2]]
tab_BMIFDmodel <- table_list[[3]]
```

```{r}
group <- c(1, NA, NA, 2, NA, NA)
quest <- c("How is your mood today?", "good or very good", "bad or very bad","What colour is your hair?", "brown", "other")
percent <- c(NA, 80, 20, NA, 50, 50)
df <- tibble(group, quest, percent)
df <- data.frame(quest, percent)

tab_BMImodel %>%
  group_rows("average BMI (decrease)", 1, 3) %>%
  group_rows("average BMI (decrease)", 4, 7)

```{r tabBMImodel}
tab_BMImodel
```

```{r tabFDmodel}
tab_FDmodel
```

```{r tabBMIFDmodel, echo = FALSE}
tab_BMIFDmodel
```
